---
title: "rm010 Exercises: Tibble Joins"
output: 
  html_document:
    keep_md: true
    theme: paper
---

## Requirements

You will need Joey's `singer` R package for this exercise. And to install that, you'll need to install `devtools`. Running this code in your console should do the trick:

```
install.packages("devtools")
devtools::install_github("JoeyBernhardt/singer")
```

Load required packages:

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(singer)
knitr::opts_chunk$set(fig.width=4, fig.height=3, warning = FALSE, fig.align = "center")
```

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

## Exercise 1: `singer`

The package `singer` comes with two smallish data frames about songs. Let's take a look at them (after minor modifications by renaming and shuffling):

```{r}
(time <- as_tibble(songs) %>% 
   rename(song = title))
```


```{r}
(album <- as_tibble(locations) %>% 
   select(title, everything()) %>% 
   rename(album = release,
          song  = title))
```

Song, artist_name, city, album

1. We really care about the songs in `time`. But, which of those songs do we know its corresponding album?

```{r}
time %>% 
  semi_join(album, by = c("song", "artist_name") )

time %>% 
  inner_join(album, by = c("song", "artist_name") )
```

outer_join can be either left or right join, or full join. 
Do we want a mutating or filtering join?
FILTERING because we really care in songs in time tibble, but which corresponds with album.
Which do we have information about album and city.
semi_join is the filtering join.
thing that identifies an observation= by song.
2 artisits, same song name? maybe put both song and artist name.
inner_join= song, artist name, year, city and album are all joined; mutating join because you added the columns, but removed all rows in time tibble that did not appear in album (don't want that)

2. Go ahead and add the corresponding albums to the `time` tibble, being sure to preserve rows even if album info is not readily available.

```{r}
time %>% 
  left_join(album, by = c("song","artist_name")) %>% 
```
Preserve everything n base tibble, but add info from another if it is available.
Still have same number of rows 

3. Which songs do we have "year", but not album info?

```{r}
time %>% 
  semi_join(album, by = "song")
```

4. Which artists are in `time`, but not in `album`?

```{r}
time %>% 
  anti_join(album, by = "artist_name")
```


5. You've come across these two tibbles, and just wish all the info was available in one tibble. What would you do?

```{r}
time %>% 
  full_join(album, by = c("song", "artist_name"))
```


## Exercise 2: LOTR

Load in the three Lord of the Rings tibbles that we saw last time:

```{r}
fell <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Fellowship_Of_The_Ring.csv")
ttow <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Two_Towers.csv")
retk <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Return_Of_The_King.csv")
```

1. Combine these into a single tibble.

```{r}
bind_rows(fell, ttow, retk)
```

Bind rows, taking tibbles and stacking them ontop of each other.

2. Which races are present in "The Fellowship of the Ring" (`fell`), but not in any of the other ones?

```{r}
fell %>% 
  anti_join(ttow, by = "Race") %>% 
  anti_join(retk, by = "Race")
```

Anti-join, specifying column that indicates the observation, how are we identifying the observation, remove everything that doesnt appear.


## Exercise 3: Set Operations

Let's use three set functions: `intersect`, `union` and `setdiff`. We'll work with two toy tibbles named `y` and `z`, similar to Data Wrangling Cheatsheet

```{r}
(y <-  tibble(x1 = LETTERS[1:3], x2 = 1:3))
```

```{r}
(z <- tibble(x1 = c("B", "C", "D"), x2 = 2:4))
```

1. Rows that appear in both `y` and `z`

```{r}
intersect(y, z)
inner_join(y,z) 
```

Same column names in both cases so easier to do intersect.
Take rows that are common. 
Innerjoin also works, take everything that is in common and put together. 

2. You collected the data in `y` on Day 1, and `z` in Day 2. Make a data set to reflect that.

```{r}
union(
  mutate(y, day = "Day 1"),
  mutate(z, day = "Day 2")
)
```

Also use bind_rows, a bit more robust, 2 tibbles and don't want the option for rows to be removed. Same rows on day one and two, want that to be reflected in the data. 

3. The rows contained in `z` are bad! Remove those rows from `y`.

```{r}
setdiff(y, z)
antijoin(y, z)
```

y is the base tibble, and don't want to include the observations that appear in z. 
Taking z and subtracting out y.
Can also use antijoin(z,y). 
Both work is because we have the same columns, if they were different variables in z and y then would have to use antijoin.